\subsection{Summary}\label{subsec:summary}
Pending Text

\subsection{Future Work}\label{subsec:future-work}
\begin{itemize}
    \item \textbf{Geodict Coordinates} Incorporating geographic data to identify areas with high demand and low supply.
    \item \textbf{Temporal Analysis:} Identifying peak hours and periods of high demand to inform redistribution strategies.
    \item \textbf{Predictive Modelling:} Developing models to forecast demand and supply, enabling proactive redistribution.
\end{itemize}
\begin{thebibliography}{24}

    \bibitem{TechProjectSourceCode}
    D. Pessoa, ``TechProject1: Source Code for Bike Sharing System Analysis,'' 2024.
    [Online].
    Available: \url{https://github.com/diogo-pessoa/TechProject1}.
    [Accessed: 07-02-2024].

    \bibitem{DivvyData}
    DivvyData, ``Bike ride-sharing data,'' 2023.
    [Online].
    Available: \url{https://divvybikes.com/system-data}

    \bibitem{DataSource}
    DivvyData Dataset, ``Index file for historic data,'' 2023.
    [Online].
    Available: \url{https://divvy-tripdata.s3.amazonaws.com/index.html}

    \bibitem{2016}
    J. Zhang, X. Pan, M. Li, and P. S. Yu, ``Bicycle-Sharing System Analysis and Trip Prediction,'' in \textit{2016 17th IEEE International Conference on Mobile Data Management (MDM)}, vol.
    1, pp.
    174-179, 2016.
    DOI: \href{https://doi.org/10.1109/MDM.2016.35}{10.1109/MDM.2016.35}

    \bibitem{2015}
    Y. Li, Y. Zheng, H. Zhang, and L. Chen, ``Traffic prediction in a bike-sharing system,'' \textit{Proceedings of the 23rd SIGSPATIAL International Conference on Advances in Geographic Information Systems}, SIGSPATIAL '15, New York, NY, USA: Association for Computing Machinery, 2015, Art.
    no.
    33.
    DOI: \href{https://doi.org/10.1145/2820783.2820837}{10.1145/2820783.2820837}

    \bibitem{s18020512}
    F. Chiariotti, C. Pielli, A. Zanella, and M. Zorzi, ``A Dynamic Approach to Rebalancing Bike-Sharing Systems,'' \textit{Sensors}, vol.
    18, no.
            2, Art.
            no.
            512, 2018.
            DOI: \href{https://www.mdpi.com/1424-8220/18/2/512}{10.3390/s18020512}

\end{thebibliography}
